---
title: rCharts + dimple | Interactive Legend
author: Timely Portfolio
github: {user: timelyportfolio, repo: rCharts_dimple, branch: "gh-pages"}
framework: bootstrap
mode: selfcontained
ext_widgets: {rCharts: "libraries/dimple"}
highlighter: prettify
hitheme: twitter-bootstrap
---

<style>
.container { width: 850px;}
</style>

# rCharts + dimple | Interactive Legend

There are plans in [`rCharts`](http://rcharts.io/site) future to provide support for partial layouts similar to [`slidify`](http://slidify.org).  For now though, there are some other options to adding script/html to customize your chart.  A good reference example comes from [`dimplejs`](http://dimplejs.org) new [interactive legends example](http://dimplejs.org/advanced_examples_viewer.html?id=advanced_interactive_legends).  In this case a small bit of javascript is added to the base chart to hide/show scatter points when the corresponding legend item is clicked.

Currently, I think the best method of adding script is through use of `slidify`.  Since `slidify` is authored by the same author as `rCharts`, they play very nicely together.  To run this reproducible example on your own, just download [this .Rmd file](http://timelyportfolio.github.io/rCharts_dimple/example48_interactive_legend.Rmd) and `slidify` it.

If you look at the code, you can see that I grab the data and create an rChart as normal.  Then I simply take advantage of markdown's script support to copy/paste the script inside of the `<script>...</script>` tags.  I only made one change to the script, since the rCharts dimple implementation uses `l` rather than `mylegend` to refer to the legend.

Other options to customize would be to create a custom html template that adds the script.  For an example of this option, see this [rCharts d3 horizon chart tutorial](http://timelyportfolio.github.io/rCharts_d3_horizon/).  If you plan to use the html inline, you could use `myrChart$html()` and then append the script before including in your finale page.

```{r echo =F, warning=F, message=F, error=F}
opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, tidy = FALSE, cache = FALSE,results = 'asis' )
```

### Get Our Data
```{r}
data <- read.delim(
  "http://pmsi-alignalytics.github.io/dimple/data/example_data.tsv"
)
### eliminate . to avoid confusion in javascript
colnames(data) <- gsub("[.]","",colnames(data))
```

### Draw Our Chart
```{r results="asis"}
require(rCharts)
d1 <- dPlot(
  SalesValue~Price,
  groups = c("SKU", "Channel", "Owner"),
  data = subset(data, Date == "01/12/2012"),
  type = "bubble",
  height = 400,
  width = 590,
  bounds = list(x=60, y=30, width=420, height=330)
)
d1$xAxis( type = "addMeasureAxis" )
d1$yAxis( type = "addMeasureAxis" )
d1$legend(
  x = 530,
  y = 100,
  width = 60,
  height = 300,
  horizontalAlign = "right"
)
d1$display("inline")
```

### Now Add Script to Make Legend Interactive

Although you won't know it unless you look at the source, I duplicate the script so that it shows in the tutorial.  This is unnecessary unless you want the script to be visible on the rendered html page.

```
<script>
        // This is a critical step.  By doing this we orphan the legend. This
        // means it will not respond to graph updates.  Without this the legend
        // will redraw when the chart refreshes removing the unchecked item and
        // also dropping the events we define below.
        myChart.legends = [];

        // This block simply adds the legend title. I put it into a d3 data
        // object to split it onto 2 lines.  This technique works with any
        // number of lines, it isn't dimple specific.
        svg.selectAll("title_text")
          .data(["Click legend to","show/hide owners:"])
          .enter()
          .append("text")
            .attr("x", 499)
            .attr("y", function (d, i) { return 90 + i * 14; })
            .style("font-family", "sans-serif")
            .style("font-size", "10px")
            .style("color", "Black")
            .text(function (d) { return d; });

        // Get a unique list of Owner values to use when filtering
        var filterValues = dimple.getUniqueValues(data, "Owner");
        // Get all the rectangles from our now orphaned legend
        l.shapes.selectAll("rect")
          // Add a click event to each rectangle
          .on("click", function (e) {
            // This indicates whether the item is already visible or not
            var hide = false;
            var newFilters = [];
            // If the filters contain the clicked shape hide it
            filterValues.forEach(function (f) {
              if (f === e.aggField.slice(-1)[0]) {
                hide = true;
              } else {
                newFilters.push(f);
              }
            });
            // Hide the shape or show it
            if (hide) {
              d3.select(this).style("opacity", 0.2);
            } else {
              newFilters.push(e.aggField.slice(-1)[0]);
              d3.select(this).style("opacity", 0.8);
            }
            // Update the filters
            filterValues = newFilters;
            // Filter the data
            myChart.data = dimple.filterData(data, "Owner", filterValues);
            // Passing a duration parameter makes the chart animate. Without
            // it there is no transition
            myChart.draw(800);
          });
  </script>
```

<script>
        // This is a critical step.  By doing this we orphan the legend. This
        // means it will not respond to graph updates.  Without this the legend
        // will redraw when the chart refreshes removing the unchecked item and
        // also dropping the events we define below.
        myChart.legends = [];

        // This block simply adds the legend title. I put it into a d3 data
        // object to split it onto 2 lines.  This technique works with any
        // number of lines, it isn't dimple specific.
        svg.selectAll("title_text")
          .data(["Click legend to","show/hide owners:"])
          .enter()
          .append("text")
            .attr("x", 499)
            .attr("y", function (d, i) { return 90 + i * 14; })
            .style("font-family", "sans-serif")
            .style("font-size", "10px")
            .style("color", "Black")
            .text(function (d) { return d; });

        // Get a unique list of Owner values to use when filtering
        var filterValues = dimple.getUniqueValues(data, "Owner");
        // Get all the rectangles from our now orphaned legend
        l.shapes.selectAll("rect")
          // Add a click event to each rectangle
          .on("click", function (e) {
            // This indicates whether the item is already visible or not
            var hide = false;
            var newFilters = [];
            // If the filters contain the clicked shape hide it
            filterValues.forEach(function (f) {
              if (f === e.aggField.slice(-1)[0]) {
                hide = true;
              } else {
                newFilters.push(f);
              }
            });
            // Hide the shape or show it
            if (hide) {
              d3.select(this).style("opacity", 0.2);
            } else {
              newFilters.push(e.aggField.slice(-1)[0]);
              d3.select(this).style("opacity", 0.8);
            }
            // Update the filters
            filterValues = newFilters;
            // Filter the data
            myChart.data = dimple.filterData(data, "Owner", filterValues);
            // Passing a duration parameter makes the chart animate. Without
            // it there is no transition
            myChart.draw(800);
          });
  </script>

### Thanks
- John Kiernander for [dimplejs](http://dimplejs.org)
- Ramnath Vaidyanathan for [slidify](http://slidify.org) and [rCharts](http://rcharts.io/site)
- Mike Bostock for his genius, especially [d3js](http://d3js.org)
