---
title: rCharts + dimple | Slopegraph
author: Timely Portfolio
github: {user: timelyportfolio, repo: rCharts_dimple, branch: "gh-pages"}
framework: bootstrap
mode: selfcontained
highlighter: prettify
hitheme: twitter-bootstrap
---
  
<style>
  .container { width: 850px;}
</style>

## rCharts + dimple | Slopegraph

James Keirstead implemented [slopegraphs in R](http://www.jameskeirstead.ca/blog/slopegraphs-in-r/) based on the the very thorough post from [Charlie Park](http://charliepark.org/slopegraphs/#the_tablegraphic).  I couldn't resist trying to do something similar with [rCharts](http://rcharts.io) and [dimplejs](http://dimplejs.org).


```{r echo =F, warning=F, message=F, error=F}
require(knitr)
opts_chunk$set(
  message = FALSE
  , warning = FALSE
  , error = FALSE
  , tidy = FALSE
  , cache = FALSE
  ,results = 'asis'
)
```

```{r echo = FALSE}
#use rCharts and dimple to kind of make a slope graph
#reference: http://www.jameskeirstead.ca/blog/slopegraphs-in-r/

#another very good reference on slopegraph
#http://charliepark.org/slopegraphs/#the_tablegraphic
#http://charliepark.org/a-slopegraph-update/

data <- read.csv(
  "https://raw.github.com/jkeirstead/r-slopegraph/master/cancer_survival_rates.csv"
)

library(rCharts)
library(jsonlite)
```

```{r echo=FALSE}
#modify show without reinstall to fix the height
Dimple2 <- setRefClass(
  "Dimple2",
  contains = "Dimple",
  methods = list(
    initialize = function(){
      callSuper(); 
    },
    #override the show here
    show = function (mode_ = NULL, ...) 
    {
        mode_ = getMode(mode_)
        switch(mode_, static = {
            viewer = getOption("viewer")
            if (!grepl("^http", LIB$url) && !is.null(viewer)) {
                temp_dir = tempfile(pattern = "rCharts")
                dir.create(temp_dir)
                suppressMessages(copy_dir_(LIB$url, file.path(temp_dir, 
                    LIB$name)))
                tf <- file.path(temp_dir, "index.html")
                writeLines(.self$render(..., static = F), tf)
                viewer(tf)
            } else {
                writeLines(.self$render(..., static = T), tf <- tempfile(fileext = ".html"))
                if (!is.null(viewer)) {
                    viewer(tf)
                } else {
                    browseURL(tf)
                }
            }
        }, server = {
            shiny_copy = .self$copy()
            shiny_copy$params$dom = "show"
            assign(".rChart_object", shiny_copy, envir = .GlobalEnv)
            if (packageVersion("shiny") > 0.7) {
                brwsr <- getOption("viewer", interactive())
            } else {
                brwsr <- getOption("shiny.launch.browser", interactive())
            }
            shiny::runApp(file.path(system.file(package = "rCharts"), 
                "shiny"), launch.browser = brwsr)
        }, inline = {
            add_ext_widgets(lib)
            return(.self$print(...))
        }, iframe = {
            chunk_opts_ = opts_current$get()
            file_ = knitr:::fig_path(".html", chunk_opts_)
            if (!file.exists(dirname(file_))) {
                dir.create(dirname(file_))
            }
            cdn = !(chunk_opts_$rcharts %?=% "draft")
            .self$save(file_, cdn = cdn)
            writeLines(c("<iframe src='", file_, "' scrolling='no' seamless", 
                paste("class='rChart", lib, "'"), "id=iframe-", params$dom, 
                "></iframe>", "<style>iframe.rChart{ width: 100%; height: 400px;}</style>"))
            return(invisible())
        }, iframesrc = {
            writeLines(
              c(
                "<iframe srcdoc='", htmlspecialchars(.self$render(...)), 
                "' scrolling='no' seamless class='rChart ",
                lib, 
                " ' height = ",
                paste0(params$height,"px"),
                " ' width = ",
                paste0(params$width,"px"),                
                paste0("id='iframe-", params$dom, "'>"),
                "</iframe>"
              )
            )
            return(invisible())
        })
    }
))

dPlot2 <- function(x, data, ...){
  myChart <- Dimple2$new()
  myChart$getChartParams(x, data, ...)
  #set these to dimple since use all of its config.yml and layouts
  #dimple2 will not exist
  myChart$setLib("dimple")
  #make this blank by default
  #not sure why this does not happen when overriding refclass
  myChart$setTemplate(afterScript="<script></script>")
  return(myChart$copy())
}
```
<h4>First Try | Out of the Box</h4>
A slopegraph most closely resembles a line graph.  Let's plot the data in a simple line graph using `type = 'line'` with dimple.

```{r}
d1 <- dPlot2(  #dPlot2 for minor internal change to iframesrc
  value ~ year,
  groups = "group",
  data = data,
  type = "line",
  height = 800,
  width = 550,
  bounds = list(x=200,y=30,height=700,width=300)
)
d1$show("iframesrc")
```

<h4>With Some Javascript Adjustments</h4>
Let's transform this simple line with some javascript.  I'll incorporate this [nice labelling example](http://bl.ocks.org/syntagmatic/4053096) to help us label our y axis.

```{r}
#with improvements from afterScript template
d1$setTemplate(
  afterScript = "
<script>
  //axes adjustments for slopegraph
  
  //based on template myChart.axes[2] should be y
  //but just to make sure do map
  var ySlope = myChart.axes.filter(function(axis){return axis.position==='y'})[0];
  var xSlope = myChart.axes.filter(function(axis){return axis.position==='x'})[0]
  
  //remove axis labels if desired
  ySlope.shapes.remove();
  //remove gridlines
  ySlope.gridlineShapes.remove();
  //remove axis title
  ySlope.titleShape.remove();
  
  var slopelabels = d3.select('#'+opts.id).select('svg').select('g').append('g')
   .attr('class','slopelabels')
  
  //get unique values for groups in data
  //note will only work with one group level
  var firstPoints = data.filter(function(d){
    return d[opts.x] == myChart.axes[0]._draw.scale().domain()[0];
  });
  
  slopelabels.selectAll('text')
    .data(firstPoints)
    .enter()
    .append('text')
    .attr('class','labels')
    //.attr('x', function(d){
    //  return xSlope._scale(d[opts.x])
    //})
    .attr('x',d3.select('.axis').select('.tick text').attr('x') - 20)
    .attr('y', function(d){
      return ySlope._scale(d[opts.y])
    })
    .attr('dy','0.2em')
    .attr('transform',d3.select('.axis').select('.tick').attr('transform'))
    .attr('fill',function(d)  {
      return myChart._assignedColors[d[opts.groups]].fill
    })
    //.attr('stroke',function(d){return myChart._assignedColors[d[opts.groups]].stroke})
    .attr('opacity',function(d){return myChart._assignedColors[d[opts.groups]].opacity})
    .style('text-anchor','end')
    .text(function(d){
      return d[opts.groups]
    });
  
    // constraint relaxation on labels
    // from http://bl.ocks.org/syntagmatic/4053096
    //add y for each of these to use code as is
    firstPoints.forEach(function(d){
      d.y = ySlope._scale(d[opts.y]);
    })
  
    var alpha = 0.5;
    var spacing = 12;
    function relax() {
      var again = false;
      firstPoints.forEach(function(a,i) {
        firstPoints.slice(i+1).forEach(function(b) {
          var dy = a.y - b.y;
          if (Math.abs(dy) < spacing) {
            again = true;
            var sign = dy > 0 ? 1 : -1;
            a.y += sign*alpha;
            b.y -= sign*alpha;
          }
        });
      });
      d3.selectAll('.labels')
        .attr('y', function(d) {
        return d.y;
      });
      if (again) setTimeout(relax,20);
    };
    
    relax();  
    
    
    //add numbers to each point
    var pointtext = d3.select('#'+opts.id).select('svg').select('g').append('g')
      .attr('class','pointtext')
    pointtext.selectAll('text')
      .data(data)
      .enter()
      .append('g')
      .attr('transform',function(d){
        return d3.select(d3.select('.axis').selectAll('.tick')[0].filter(function(dd){
          return d3.select(dd).datum() == d[opts.x]
        })[0]).attr('transform')
      })
      .append('text')
      .attr('x',function(d){
        return d3.select('.axis').select('.tick text').attr('x')
      })
      .attr('y',function(d){
        return ySlope._scale(d[opts.y])})
      .attr('dy','0.2em')
      //.attr('fill',function(d){return myChart._assignedColors[d[opts.groups]].fill})
      //.attr('stroke',function(d){return myChart._assignedColors[d[opts.groups]].stroke})
      //.attr('opacity',function(d){return myChart._assignedColors[d[opts.groups]].opacity})
      .attr('text-anchor','middle')
      .style('font-size','12')
      .style('pointer-events','none')
      .text(function(d){
        return d[opts.y]
      })
</script>"
)
d1$show("iframesrc")
```
