---
title: rCharts, dimple, and time series
subtitle: Some Experiments Using US Treasury Yield Data
author: Timely Portfolio
github: {user: timelyportfolio, repo: rCharts_dimple, branch: "gh-pages"}
framework: bootstrap
mode: selfcontained
ext_widgets: {rCharts: "libraries/dimple"}
highlighter: prettify
hitheme: twitter-bootstrap
---

<style>
/*
path.line {
  fill: none;
  stroke: #666;
  stroke-width: 1.5px;
}
*/
.axis {
  shape-rendering: crispEdges;
}

.axis text {
  font: 10px sans-serif;
  margin: 0;
}

.x.axis line {
  stroke: #000;
}

.x.axis path {
  display: none;
}
</style>

```{r echo=F, warning= F, message=F}
opts_chunk$set(message = FALSE, warning = FALSE, error = FALSE, tidy = FALSE, cache = FALSE,results = 'asis' )
```

Even this early in it release, [Dimple](http://dimplejs.org) is such a powerful and promising [d3](http://d3js.org) library that we decided to quickly make it available in [rCharts](http://rcharts.io/site).  The first test was to recreate all (48) of the examples provided by [dimplejs](http://dimplejs.org/examples_index.html) with `rCharts`.  Once we were satisfied with those, we demonstrated fairly good coverage of the `dimplejs` API (easily one of the best [documented](https://github.com/PMSI-AlignAlytics/dimple/wiki)).  We are aware of a couple missing items, but I really wanted to throw some real financial time series at it to see how well it might work in my real life as a portfolio manager.

Since bonds are a hot topic now, I thought some US Treasury Yield data from the St. Louis Federal Reserve (FRED) would make a nice subject.  I will start very basic and build up to a still simple but slightly more complicated plot.  As always, we will need to grab the data with `quantmod` `getSymbols()` and then merge it into one big `ust.xts` object.

```{r}
require(quantmod)
require(rCharts)
#now get the US bonds from FRED
USbondssymbols <- paste0("DGS",c(1,2,3,5,7,10,20,30))

ust.xts <- xts()
for (i in 1:length( USbondssymbols ) ) {
  ust.xts <- merge( 
    ust.xts,
    getSymbols( 
      USbondssymbols[i], auto.assign = FALSE,src = "FRED"
    )
  )
}
```

Then we will define a little `xtsMelt` function to transform our wide data into long form.

```{r}
xtsMelt <- function(data) {
  require(reshape2)
  
  #translate xts to time series to json with date and data
  #for this behavior will be more generic than the original
  #data will not be transformed, so template.rmd will be changed to reflect
  
  
  #convert to data frame
  data.df <- data.frame(cbind(format(index(data),"%Y-%m-%d"),coredata(data)))
  colnames(data.df)[1] = "date"
  data.melt <- melt(data.df,id.vars=1,stringsAsFactors=FALSE)
  colnames(data.melt) <- c("date","indexname","value")
  #remove periods from indexnames to prevent javascript confusion
  #these . usually come from spaces in the colnames when melted
  data.melt[,"indexname"] <- apply(matrix(data.melt[,"indexname"]),2,gsub,pattern="[.]",replacement="")
  return(data.melt)
  #return(df2json(na.omit(data.melt)))
}
```

Now let's use our `xtsMelt` function on our data.  You might also notice that we use `na.omit` to remove the pesky NAs that sometimes come over in FRED data.

```{r}
ust.melt <- na.omit( xtsMelt( ust.xts["2012::",] ) )

ust.melt$date <- format(as.Date(ust.melt$date))
ust.melt$value <- as.numeric(ust.melt$value)
ust.melt$indexname <- factor(
  ust.melt$indexname, levels = colnames(ust.xts)
)
ust.melt$maturity <- as.numeric(
  substr(
    ust.melt$indexname, 4, length( ust.melt$indexname ) - 4
  )
)
ust.melt$country <- rep( "US", nrow( ust.melt ))
```

We have our data ready to plot.  I hope you are impressed with how easy it is to get an interactive `dimplejs` chart from R.  Without comments, it would all fit nicely on one line.  We can even use a `r` formula to define our 'x' and 'y'.

```{r}
#simple line chart of 10 year
d1 <- dPlot(
  value ~ date,  #or x="date", y="value"
  #dimplejs allows filtering but will lessen data to write
  #if we subset in R
  data = subset(ust.melt,maturity==10),  #get all data for 10y maturity
  type = 'line'
)
d1
```
```{r echo = F}
d1$print('chart1')
```

Uh oh, our x axis does not look too pretty.  `rCharts` is extensible and modular, so quickly we will need to move to a little more advanced concept.  `dimplejs` is built on `d3`, so let's fix with a little [help](http://bl.ocks.org/mbostock/1166403) from the `d3` master [Mike Bostock](http://bost.ocks.org/mike/).  I built a custom [layout template](http://timelyportfolio.github.io/rCharts_dimple/assets/chart_d3_dateaxis.html) to remove the `dimple` x-axis and replace with a `d3` version featuring much better date/time support.  To access it, we can set it with a little hack (I have a little inside information that this will soon become much easier).

```{r}
d1$field(
  'templates',
  modifyList(p1$templates, list(script = 'chart_d3dateaxis.html')
)
d1
```

```{r echo = F}
print()
```

```{r}
d1 <- dPlot(
  x = c("maturity","date"),
  y = "value",
  groups = 'maturity',
  data = ust.melt,
  type = 'line'
)
d1$xAxis(grouporderRule = "date")
d1$print('chart1')
```




